---
title: "Final project"
author: "Mingyang Zheng"
date: "10/24/2020"
output: html_document
---

```{r setup, include=FALSE}
## preparation work
setwd("~/Desktop/R/datasciencecoursera/testdir")
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)

#load files
pml <- read.csv("pml-training.csv")
pml_testing <- read.csv("pml-testing.csv")

#convert variable classe to factor
pml$classe <- as.factor(pml$classe )

names(pml)
```


## data cleaning process

### I first remove variables that contains NA values. I also remove the first seven variables (user names, row numbers, etc) because they has nothing to do with the outcome.Lastly, I remove variables that large have empty inputs. By the end of the data cleanning process, the dataset has 53 variables. 

```{r echo=FALSE}
pml<- pml[, colSums(is.na(pml)) == 0]
names(pml)
# remove first 7 colums because it has nothing to do with the outcome
pml = subset(pml, select = -c(1:7) )
# remove coloums that have lot of empty inputs
pml = subset(pml, select = -c(5:13) )
pml = subset(pml, select = -c(27:32) )
pml = subset(pml, select = -c(30:38) )
pml = subset(pml, select = -c(43:51) )

dim(pml)
```


## Building training and testing dataset

### I creat a training dataset and a testing dataset using a 70/30 split. 
```{r echo=FALSE}

set.seed(343434)

inTrain <- createDataPartition(y = pml$classe,
                               p = 0.7, list = F)

training <- pml[inTrain,]
testing <- pml[-inTrain,]

```


## Predict model

### I use random forest method because this random forest is versatile. I also set ntree to 250 in order to save the computational power. 

### The accuracy of the model is 0.9969, meaning that the out of sample error is 0.0031. 

```{r echo=FALSE}

model_rf <- train(classe ~ ., method="rf",ntree=250, data=training)  

model_predict <- predict(model_rf, testing)

confusionMatrix(model_predict, testing$classe)
```

### Lastly, I use the model to predict the 20 different test cases.

```{r echo=FALSE}
model_predict2 <- predict(model_rf, pml_testing)

model_predict2

```



